Vragenlijst thesis evaluatie

### 2015-06-22 Thesisfollowup/ThesisFollowUpSophieMarien.tex
nelezen verbeteringen

Thesis follow up vragen van Maart en April terug beantwoord. 
\section{Hoever sta je? Moet het werkplan aangepast worden?}
 Momenteel moet ik mijn tekst verfijnen van het hoofdstuk over worm propagatie en de related work. De nash equilibria moeten ook nog berekent worden. Na de examens ga ik terug verder werken aan het berekenen van de Nash equilibria. Ondertussen werk ik elke avond een deel van de tekst af. De hoofdstukken tot en met het Nash Equilibria zijn al volledig uitgeschreven. 
\section{Wat zijn de belangrijkste elementen uit het besluit dat je gaat schrijven in je tekst?}
In mijn besluit ga ik, als ze bestaan, kunnen vaststellen wat de nash equilibria zijn en hier conclusies kunnen uittrekken voor de verdedigings-strategie"en van de defender. 
\section{Verfijn de outline van je thesistekst. Geef per hoofdstuk de subsecties, en per subsectie het kernidee en een schatting van het aantal bladzijdes van die subsectie.}

\begin{enumerate}
\item \textbf{Introduction} 
\begin{itemize}
\item Research questions
\item Contributions and results
\item Conjunctures and open problems
\item Overview thesis
\end{itemize}
\item \textbf{General context}
\begin{itemize}
\item[2.1] What is security
\item[2.2] A brief introduction to game theory
\item[2.3] The FlipIt game
\item[2.4] Extensions on FlipIt
\end{itemize}
\item \textbf{FlipIt with propagation delay}
\begin{itemize}
\item[3.1] Introduction
\item[3.2] Difference between FlipIt with and without propagation delay
\begin{itemize}
\item Single resource
\item Only one action per player
\item Immediate effect of the move
\item Stealth character of the moves
\item Cost associated with the move
\item Strategies
\item Application: Wormlike ATP
\end{itemize}
\item[3.3] Playing periodically with propagation delay
\end{itemize}
\item \textbf{Nash Equilibria}
\begin{itemize}
\item[4.1] Determining piecewise functions $opt_{D}(\delta_{A}$
\item[4.2] Determining piecewise functions $opt_{A}(\delta_{D}$
\item[4.3] Determing the Nash equilibria
\end{itemize}
\item\textbf{ Worm propagation}
\begin{itemize}
\item[5.1] Methods of worm propagation
\item[5.2] Graph models
\item[5.3] Models for worm propagation
\item[5.4] Method for calculating propagation delay through matrices
\end{itemize}
\item \textbf{Related Work}
\item \textbf{Conclusion}
\end{enumerate}

\section{Beschrijf in 10 lijnen de demo die je bij je verdediging zult geven.}
Ik heb een theoretische thesis dus kan ik moeilijk een demo geven. Ik zal enkele voorbeelden uitwerken om mijn werkingsmethode uit te leggen.
\section{Wat is er afgesproken met je begeleider en/of promotor om op te leveren als major milestone die het praktische werk van je thesis afsluit?}
Dit is besproken met de begeleiders maar er is niets afgesproken omdat op voorhand niet kan gezegd worden of de analyse van het FlipIt spel voldoende zal zijn omdat de indruk kan ontstaan dat dit resultaat eenvoudig te verkrijgen was. Er is dus niet afgesproken wat ik moet inleveren om mijn thesis af te sluiten.
%Als ik hier niet mee akkoord dan moet ik dit ook verduidelijken, en moet ik mijn standpunt verdedigen (in mijn tekst, en op de verdediging). 
Voorlopig werk ik verder af wat ik zelf al voor ogen had; namelijk het vinden van de nash equilibria en het schrijven van een related work over worm propagatie. \\
Mijn promotor stelde het volgende: Mijn masterproef is in de eerste plaats mijn verantwoordelijkheid, ik ben verantwoordelijk voor wat ik indien en wat ik doe. De rol van mijn begeleiders en promotor is hierbij advies en richtlijnen geven.
Als ik het hiermee niet eens ben dan zorg ik voor de nodige argumenten en stel een sterke tekst op samen met een goede verdediging van mijn werk. 
\section{Evalueer de tijd die je in je thesis gestoken hebt in de voorbije 4 weken: veel te weinig / te weinig / voldoende / ruim voldoende / meer dan verwacht.}
De voorbije weken heb ik meer dan verwacht aan mijn thesis gewerkt. Momenteel ben ik al aan het aantal uren gekomen dat ik aan mijn thesis moet werken. Hierbij zullen waarschijnlijk nog eens 150 uren bijkomen. 
\section{Evalueer de vorderingen van je thesis t.o.v. het bereiken van de vooropgestelde doelstelling: ver achter op het schema / achter op het schema / goed op schema / vooruit op het schema / ver vooruit op het schema.}
Ik zit achter op schema want anders had ik al in eerste zit kunnen afgeven. De komende week probeer ik mijn Nash equilibria hoofdstuk af te werken en daarna focus ik mij op het schrijven van de hoofdstukken.
\section{Stuur de ingevulde timesheet mee als bijlage.}

>>>>

Thesis follow up vragen van Maart en April terug beantwoord. 
\section{Hoever sta je? Moet het werkplan aangepast worden?}
Het werkplan werd aangepast in functie van indiening in Augustus. Doelstellingen voor juni zijn: afwerken van de tekst van hoofdstukken 1,2, 3 en 5 en finaliseren van de formules voor de Nash evenwichten.  Momenteel moet ik nog de tekst verfijnen van hoofdstuk 5 over worm propagatie en de related work. De formules voor de nash equilibria zijn al deels ontwikkelde, maar moeten nog verder uitgewerkt worden. De week van 22 juni is gereserveerd voor de examens. Vanaf 27 juni ga ik terug verder werken aan het berekenen van de Nash equilibria.  Tussen 6 en 31 juli wordt de tekst door vrienden nagelezen op tikfouten, structuur en verstaanbaarheid. In augustus wordt de tekst van hoofdstukken 4 en 6 afgewerkt 
\section{Wat zijn de belangrijkste elementen uit het besluit dat je gaat schrijven in je tekst?}
\begin{itemize}
\item vaststellen wat de nash equilibria zijn
\item conclusies voor de verdedigingsstrategie"en van de defender
\item conclusies voor de aanvalsstrategie"en van de attacker
\item Ide"en voor verder onderzoek. 
\end{itemize}

\section{Verfijn de outline van je thesistekst. Geef per hoofdstuk de subsecties, en per subsectie het kernidee en een schatting van het aantal bladzijdes van die subsectie.}

\begin{enumerate}
\item \textbf{Introduction}(4 blz., final draft afgewerkt) 
\begin{itemize}
\item Research questions
\item Contributions and results
\item Conjunctures and open problems
\item Overview thesis
\end{itemize}
\item \textbf{General context}(11 blz., final draft afgewerkt)
\begin{itemize}
\item[2.1] What is security
\item[2.2] A brief introduction to game theory
\item[2.3] The FlipIt game
\item[2.4] Related Work on Extensions to FlipIt
\end{itemize}
\item \textbf{FlipIt with propagation delay} (13 blz., final draft afgewerkt)
\begin{itemize}
\item[3.1] Introduction
\item[3.2] Difference between FlipIt with and without propagation delay
\begin{itemize}
\item Single resource
\item Only one action per player
\item Immediate effect of the move
\item Stealth character of the moves
\item Cost associated with the move
\item Strategies
\item Application: Wormlike ATP
\end{itemize}
\item[3.3] Formalization of the periodic game with propagation delay
\end{itemize}
\item \textbf{Nash Equilibria}(formules deels uitgewerkt, geschatte lenge: 12 blz.)
\begin{itemize}
\item[4.1] Determining piecewise functions $opt_{D}(\delta_{A})$
\item[4.2] Determining piecewise functions $opt_{A}(\delta_{D})$
\item[4.3] Determining the Nash equilibria
\end{itemize}
\item\textbf{Models for the Delay}(8 blz., final draft afgewerkt)
\begin{itemize}
\item[5.1] Methods of worm propagation
\item[5.2] Graph models
\item[5.3] Models for worm propagation
\item[5.4] Method for calculating propagation delay through matrices
\end{itemize}
\item \textbf{Related Work: Alternative approaches to the security problem}  
\begin{itemize}
\item (optioneel hoofdstuk nog uit te werken in functie van beschikbare tijd, geschatte lengte: 2 blz.
\item In dit hoofdstuk zou ik alle literatuur willen bespreken die ik aan het begin van mijn thesis heb gelezen met het oog op het afbakenen van het onderwerp.  Dat zijn een hele hoop papers, maar er moet nog een goed stramien gevonden worden voor het structureren van de bespreking.
\end{itemize}
\item \textbf{Conclusion}(nog uit te werken, geschatte lengte: 2 blz.) 
\end{enumerate}

\section{Beschrijf in 10 lijnen de demo die je bij je verdediging zult geven.}
Ik heb een theoretische thesis en dus kan ik geen live demo geven. Ik ben van plan om mijn werkingsmethode uit te leggen aan de hand van enkele voorbeelden die ik in figuren zal uitwerken.
\section{Wat is er afgesproken met je begeleider en/of promotor om op te leveren als major milestone die het praktische werk van je thesis afsluit?}
Dit werd op de laatste vergadering besproken met de begeleiders. Ik stuurde na de vergadering een samenvattend verslag van de bespreking, maar de begeleiders antwoordden dat er op voorhand niet kan gezegd worden of de analyse van het FlipIt spel voldoende zal zijn omdat de indruk kan ontstaan dat dit resultaat eenvoudig te verkrijgen was. Er is dus niets afgesproken wat ik moet inleveren om mijn thesis af te sluiten.
%Als ik hier niet mee akkoord dan moet ik dit ook verduidelijken, en moet ik mijn standpunt verdedigen (in mijn tekst, en op de verdediging). 
Voorlopig werk ik verder af wat ik zelf al voor ogen had; namelijk het vinden van de nash equilibria en het schrijven van een related work over worm propagatie. \\
Ik stelde dezelfde vraag aan mijn promotor en kreeg het volgende antwoord: Mijn masterproef is in de eerste plaats mijn verantwoordelijkheid, ik ben verantwoordelijk voor wat ik indien en wat ik doe. De rol van mijn begeleiders en promotor is hierbij advies en richtlijnen geven.
Als ik het hiermee niet eens ben dan zorg ik voor de nodige argumenten en stel een sterke tekst op samen met een goede verdediging van mijn werk. 
Besluit: Er is dus vanuit promotor en werkleider geen concrete afspraak over wat ik moet inleveren om mijn thesis af te sluiten. Ik ben zelf van oordeel dat het uitwerken van de periodische strategie volstaat om tegemoet te komen aan de doelstelling van mijn masterproef en zal dit beargumenteren in de tekst van mijn masterthesis. 
\section{Evalueer de tijd die je in je thesis gestoken hebt in de voorbije 4 weken: veel te weinig / te weinig / voldoende / ruim voldoende / meer dan verwacht.}
De voorbije weken heb ik meer dan verwacht aan mijn thesis gewerkt. Momenteel heb  ik al 765 uren gewerkt. Hier zullen naar schatting nog eens 150 uren bijkomen. 
\section{Evalueer de vorderingen van je thesis t.o.v. het bereiken van de vooropgestelde doelstelling: ver achter op het schema / achter op het schema / goed op schema / vooruit op het schema / ver vooruit op het schema.}
Het geplande werk voor juni is uitgevoerd volgens schema. Tot 28 juni werk ik de tekst van hoofdstuk 5 verder af. In de week van 29 juni probeer ik mijn Nash equilibria hoofdstuk af te werken. 
\section{Stuur de ingevulde timesheet mee als bijlage.}

### 2015-06-22 doc/template/chap-4.tex
nelezen verbeteringen

\chapter{Worm Propagation }
\label{chapter4: Worm propagation}

The spreading of malware has already been extensively researched. Because of all the different types of propagation it is hard to pick out one model that can model all of them. Modelling the spread of malware depends on two great factors: the method used for the propagation and the graph where the malware will spread itself. Viruses and worms are the types of malware that are the most researched. \\
In this chapter we focus on the propagation models of worms. The propagation of a worm is not depending on the human interaction as it is with viruses. In section [] we present an overview of the most used propagation methods by worms. Next, in section [] different kind of graphs to model a computer network are introduced. Section [] lists all the kind of models to determine the time a worm needs to infect a sufficient number of nodes. These results can be used for the completion of parameter \textit{d}. Finally we introduce an easy method to calculate the delay of the propagation of a worm.

@@

\subsection{Methods of propagation}

There exist two kinds of APT's in the context of propagation. There are APT's that only launch an attack on just one target. The mechanism they use to propagate is the dropper mechanism. The dropper is the initial attack vector that compromises the system. The other kind of APT's target multiple victims. They also use a dropper mechanism and have an additional mechanism for self-propagation. It can be either the virus propagation or the worm like propagation. A virus infects one node on the network and has to wait for human interaction to spread. So the spreading speed depends on the human interaction. If an APT uses a worm propagation method it will spread by itself after it has been dropped on the network. Because it is not easy to incorporate the human factor in a spreading method we will only examine the worm propagation models for our APT. \\

Below is a list of several common propagation methods:

\begin{description}
\item Scanning methods: A worm can use distinct ways of scanning: random, localized, topological or hit-list scanning.
\begin{itemize}
\item Random scanning: Some worms propagate by using the method of random targeting IP addresses. Rate of terms of succeeded randomly chosen IP addresses is very low. An example of such a worm is `Code Red' ans 'Slammer'.
\item Localized scanning: --Localized scanning,
used by the Code Red II and Nimda worms, preferentially
scans for hosts in the ''local'' address space.--
\item Topological scanning: --Topological scanning, used by the Morris worm, relies
on the ''address'' information contained in the victim
machines to locate new targets.--
\item Hit-list scanning: With the method of hit-list scanning, before hand a short-list of vulnerable systems are made. --to speed up
the spread of worms at the initial stage. This list consists
of potentially vulnerable machines that are gathered
beforehand and targeted first when the worm
is released. An extreme case for the hit-list-scanning
worms is a flash worm, which gathers all vulnerable
machines into the list.--
\end{itemize}   
\item Routing worms: they use BGP routing tables to scan the routable address space. 3 times faster than traditional worm that uses random scanning. Spyb0t or network Bluepill. --a ''routing worm'' [20]. Zou et al. designed two types
of routing worms [20]. One type, based on Class-A
(x.0.0.0/8) address allocations, is thus called 'ClassA
routing worms.'' Such worms can reduce the scanning
space to 45.3\% of the entire IPv4 address space.
The other type, based on BGP routing tables, is thus
called''BGP routing worms'' Such worms can reduce
the scanning space to only about 28.6\% of the entire
IPv4 address space.--
\item email worms: use the email systems to propagate. The Kak worm is a Javascript computer worm that spread itself by exploiting a bug in Outlook Express.  GhostNet
\item DNS random scanning: --Another strategy that a worm
can potentially employ is DNS random scanning [5],
in which a worm uses the DNS infrastructure to locate
likely targets by guessing DNS names instead of
IP addresses. Such a worm in the IPv6 Internet is
shown to exhibit a propagation speed comparable to
that of an IPv4 random-scanning worm--
%Pikachu worm: The virus was mainly spread through Microsoft Outlook email attachments. The email containing the attached virus propagated through infected users by sending itself to all contacts in the user's Outlook address book. [5]
\item self stopping worm: reduces speed to avoid detection. Atak worm or self stopping worm
\item network sharing Shamoon
\item shared files or spreading through SMB (shared message block or  Common Internet File System (CIFS)) Flame
\item zero-day vulnerabilities
\end{description}

--dit stuk tekst overbodig--
The start of such a propagation can be different. Either it is a targeted attack or a random attack. For example a usb stick or other shared resources can be used to start the spreading. It can be given to some special person or it can just be left behind in a parking lot. This can still be a targeted attack if it is the parking lot of a targeted company but there is more insecurity of where the usb stick will end. Another way is by sending an email. It can also start from one computer that is connected directly to the internet. Downloading malicious programs.

@@

%So the chance that a resource is infected after time t is the chance that the resource is infected by all the previous infections and that the defender has not flipped the resource.
source: \url{http://en.wikipedia.org/wiki/Adjacency_matrix} \\
We model the network through an undirected Graph $G = < V, E> $ where $|V|$ denotes the number of resources in the network and $|E|$ the number of connections. We can convert this to a adjacent matrix where we can represent which vertices of the graph are neighbours of other vertices. \\
For our graph we have an $|V| \times |V|$ matrix with on every entry $a_{ij}$ a 1 as value if there is a connection between node $V_{i}$ and $V_{j}$ and with zeros its diagonal. Because our graph is undirected we have a symmetric matrix.

\textit{"If \textit{A} is the adjacency matrix of the directed or undirected graph \textit{G}, then the matrix $A^{n}$ (i.e., the matrix product of n copies of \textit{A}) has an interesting interpretation: the entry in row i and column j gives the number of (directed or undirected) walks of length n from vertex i to vertex j. If n is the smallest nonnegative integer, such that for all i ,j , the (i,j)-entry of $A^{n} > 0$, then n is the distance between vertex i and vertex j."} [Wikipedia]

@@

  \path (A) edge              node {} (B)
            edge              node {} (D)
        (B) edge              node {} (A)
        	edge			  node {} (C)
        (C) edge              node {} (B)
            edge 			  node {} (D)
            edge			  node {} (E)
            edge			  node {} (F)
        (D) edge 			  node {} (C)
            edge              node {} (A)
        (E) edge 			  node {} (C)
    	(F)	edge			  node {} (C);
\end{tikzpicture}
\\
%\[
%\begin{bmatrix}
%    x_{11}       & x_{12} & x_{13} & \dots & x_{1n} \\
%    x_{21}       & x_{22} & x_{23} & \dots & x_{2n} \\
%    \hdotsfor{5} \\
%    x_{d1}       & x_{d2} & x_{d3} & \dots & x_{dn}
%\end{bmatrix}
%=
%\begin{bmatrix}
%    x_{11} & x_{12} & x_{13} & \dots  & x_{1n} \\
%    x_{21} & x_{22} & x_{23} & \dots  & x_{2n} \\
%    \vdots & \vdots & \vdots & \ddots & \vdots \\
%    x_{d1} & x_{d2} & x_{d3} & \dots  & x_{dn}
%\end{bmatrix}
%\] 
The adjacent matrix becomes this matrix $[A]$: \\

@@

Matrix $A \times A = A^{2}$ becomes the matrix with the number of paths with 2 steps from $N_{i}$ to $N_{j}$: We denote this matrix as matrix \textit{[B]}\\

@@

Matrix $A^{2} \times A = A^{3}$ becomes the matrix with the number of paths with 3 steps from $N_{i}$ to $N_{j}$: We denote this matrix as matrix \textit{[C]}\\

$
\bordermatrix{
         & N_1		& N_2	& N_3	& N_4 	& N_5	&N_6     \cr
    N_1   & 0		& 4		& 4		& 0		& 2		& 2	     \cr
    N_2   & 4		& 0		& 0		& 6		& 0		& 0	     \cr
    N_3   & 4		& 0		& 0		& 6		& 0		& 0	     \cr
    N_4   & 0		& 6		& 6		& 0		& 4		& 4	     \cr
	N_5   & 2		& 0		& 0		& 4		& 0		& 0	     \cr
	N_6   & 2		& 0		& 0		& 4		& 0		& 0	     \cr
}$ 
\\

So for $A^{N}$ every $a_{ij}$ entry gives the number of paths with N steps from $N_{i}$ to $N_{j}$.\\

With this knowledge we can calculate in how many steps a node is infected. $A$ calculates which nodes are infected after 1 step, $A^{N}$ calculates which nodes are infected in N steps.. So if we want to know how many nodes are infected after 3 steps we have to add every matrix $(A + A^{2} + A^{3}) $ and see which entry is a non zero entry.

>>>>

\chapter{Models for the Delay }
\label{chapter4: Worm propagation}

\todo{nog uitleggen wat dit hoofdstuk hier staat te doen.
The formalisation of the FlipIt game with delay relies on some value "d" that represents the time needed to infect a sufficient number of nodes in a network after the initial infection. This chapter provides the reader with more insight on how to calculate the value of this parameter d. 
}
The  spreading of malware has already been extensively researched. Because of the many different types of propagation it is hard to defined a single model that can model all of them. Modelling the spread of malware depends on two key factors: the method used for the propagation and the graph where the malware will spread itself. \\
Viruses and worms are the types of malware that are the most researched. Since the spreading of viruses requires human interaction, their propagation delay is depending on (hard to predict) human behaviour. Worms on the other hand, spread without human intervention, and are their propagation is therefore easier to model. Since the purpose of this chapter is only to illustrate how a delay can be calculated, we will limit the presentation to propagation models of worms . Section []  presents an overview of the most frequently used propagation methods by worms. Next, section [] introduces  different kind of graphs to model a computer network are introduced. Section [] presents    different kinds of models to determine the time a worm needs to infect a sufficient number of nodes. These results can be used for the completion of parameter \textit{d}. Finally we introduce an easy method to calculate the delay of the propagation of a worm.

@@

\subsection {Methods of propagation}

In the context of malware propagation, there exists two kinds of APT's . First, there are APTs that launch an attack on just one target node of a network. The mechanism these APT's use to propagate malware is the dropper mechanism. The dropper is the initial attack vector that compromises the single targeted node of the system. The second kind of APTs target multiple victims. These APTs also use a dropper mechanism but have an additional mechanism for self-propagation in order to propagate themselves to the multiple targeted nodes of the system. This mechanism can either be a virus propagation or a worm propagation. A virus infects one node on the network and has to wait for human interaction to spread. So the spreading speed depends on the human interaction, and modeling virus propagation therefore requires modeling human behavior. If an APT uses a worm propagation method it will spread by itself after it has been dropped on the network. Given the additional complexity of incorporating the human factor in a spreading method and that this chapter is merely meant as illustration on calculating the delay for use in the game theoretic approach, we will limit the presentation to worm propagation models for APT's. \\

Infection by worms start with dropping the work on the network. Different dropping mechanisms can be used, whereby the initial attack can be random or targeted at a specific node. Frequently used mechanisms are USB sticks (given to a specific person or left behind for pick up by a random person), email, or malicious software. More important for determining our delay is the propagation strategy. Propagation is performed by determining the next nodes to spread the worm to. The following are common propagation methods: 
\begin{description}
\item Scanning methods: A worm tries to guess (scan) the address of potential target nodes to infect. It can use distinct ways of scanning: random, localized, topological or hit-list scanning.
\begin{itemize}
\item Random scanning: Some worms propagate by using the method of random targeting IP addresses. The rate of success for randomly chosen IP addresses is very low. Example of such worms are `Code Red' and 'Slammer '.
\item Localized scanning: --Localized scanning,
used by the Code Red II and Nimda worms, preferentially
scans for hosts in the ''local'' address space.--
\item Topological scanning: --Topological scanning, used by the Morris worm, relies
on the ''address'' information contained in the victim
machines to locate new targets.--
\item Hit-list scanning: With the method of hit-list scanning, a short-list of vulnerable systems is made beforehand. --to speed up
the spread of worms at the initial stage. This list consists
of potentially vulnerable machines that are gathered
beforehand and targeted first when the worm
is released. An extreme case for the hit-list-scanning
worms is a flash worm, which gathers all vulnerable
machines into the list.--
\end{itemize}   
\item Routing worms: they use BGP routing tables to scan the routable address space. 3 times faster than traditional worm that uses random scanning. Spyb0t or network Bluepill. --a ''routing worm'' [20]. Zou et al. designed two types
of routing worms [20]. One type, based on Class-A
(x.0.0.0/8) address allocations, is thus called 'ClassA
routing worms.'' Such worms can reduce the scanning
space to 45.3\% of the entire IPv4 address space.
The other type, based on BGP routing tables, is thus
called''BGP routing worms'' Such worms can reduce
the scanning space to only about 28.6\% of the entire
IPv4 address space.--
\item email worms: use the email systems to propagate. The Kak worm is a Javascript computer worm that spread itself by exploiting a bug in Outlook Express.  GhostNet
\item DNS random scanning: --Another strategy that a worm
can potentially employ is DNS random scanning [5],
in which a worm uses the DNS infrastructure to locate
likely targets by guessing DNS names instead of
IP addresses. Such a worm in the IPv6 Internet is
shown to exhibit a propagation speed comparable to
that of an IPv4 random-scanning worm--
%Pikachu worm: The virus was mainly spread through Microsoft Outlook email attachments. The email containing the attached virus propagated through infected users by sending itself to all contacts in the user's Outlook address book. [5]
\item self stopping worm: reduces speed to avoid detection. Atak worm or self stopping worm
\item network sharing Shamoon
\item shared files or spreading through SMB (shared message block or  Common Internet File System (CIFS)) Flame
\item zero-day vulnerabilities
\end{description}

--dit stuk tekst overbodig--

@@

%So the chance that a resource is infected after time t is the chance that the resource is infected by all the previous infections and that the defender has not flipped the resource.
source: \url{http://en.wikipedia.org/wiki/Adjacency_matrix} \\
We model the network through an undirected Graph $G = < V, E> $ where $|V|$ denotes the number of resources in the network and $|E|$ the number of connections. We can convert this to an adjacency matrix which represents which vertices of the graph are neighbors of other vertices. \\
The graph is represented as a $|V| \times |V|$ matrix with for every entry $a_{ij}$ a 1 as value if there is a connection between node $V_{i}$ and $V_{j}$ and 0 otherwise, and with 0's for every $a_{ii}$. Because the graph is undirected we have a symmetric matrix. 
Adjacency matrices have many interesting applications, amongst which calculating the paths between vertices:
\textit{"If \textit{A} is the adjacency matrix of the directed or undirected graph \textit{G}, then the matrix $A^{n}$ (i.e., the matrix product of n copies of \textit{A}) has an interesting interpretation: the entry in row i and column j gives the number of (directed or undirected) walks of length n from vertex i to vertex j. If n is the smallest nonnegative integer, such that for all i ,j , the (i,j)-entry of $A^{n} > 0$, then n is the distance between vertex i and vertex j."} [Wikipedia]

@@

  \path (A) edge              node {} (B)
            edge              node {} (D)
        (B) edge              node {} (A)
        	edge			  node {} (C)
        (C) edge              node {} (B)
            edge 			  node {} (D)
            edge			  node {} (E)
            edge			  node {} (F)
        (D) edge 			  node {} (C)
            edge              node {} (A)
        (E) edge 			  node {} (C)
    	(F)	edge			  node {} (C);
\end{tikzpicture}
\\
%\[
%\begin{bmatrix}
%    x_{11}       & x_{12} & x_{13} & \dots & x_{1n} \\
%    x_{21}       & x_{22} & x_{23} & \dots & x_{2n} \\
%    \hdotsfor{5} \\
%    x_{d1}       & x_{d2} & x_{d3} & \dots & x_{dn}
%\end{bmatrix}
%=
%\begin{bmatrix}
%    x_{11} & x_{12} & x_{13} & \dots  & x_{1n} \\
%    x_{21} & x_{22} & x_{23} & \dots  & x_{2n} \\
%    \vdots & \vdots & \vdots & \ddots & \vdots \\
%    x_{d1} & x_{d2} & x_{d3} & \dots  & x_{dn}
%\end{bmatrix}
%\] 
Assuming a network like in Fig, the corresponding adjacency matrix is the  matrix $[A]$: \\

@@

In matrix $A \times A = A^{2}$, each entry represents the number of 2 step paths from $N_{i}$ to $N_{j}$: We denote this matrix as matrix \textit{[B]}\\

@@

Likewise, in matrix $A^{2} \times A = A^{3}$ each entry represents the number of paths with 3 steps from $N_{i}$ to $N_{j}$: We denote this matrix as matrix \textit{[C]}\\

$
\bordermatrix{
         & N_1		& N_2	& N_3	& N_4 	& N_5	&N_6     \cr
    N_1   & 0		& 4		& 4		& 0		& 2		& 2	     \cr
    N_2   & 4		& 0		& 0		& 6		& 0		& 0	     \cr
    N_3   & 4		& 0		& 0		& 6		& 0		& 0	     \cr
    N_4   & 0		& 6		& 6		& 0		& 4		& 4	     \cr
	N_5   & 2		& 0		& 0		& 4		& 0		& 0	     \cr
	N_6   & 2		& 0		& 0		& 4		& 0		& 0	     \cr
}$

~~\\
So, in $A^{N}$ every $a_{ij}$ entry gives the number of paths with N steps from $N_{i}$ to $N_{j}$.\\

Using this knowledge we can calculate in how many steps a node is infected: 
The infection vector or start vector I0 of lenght $|V|$ indicates which node is infected and which not. Multiplying the infection vector with $A$ results in a vector I1, which represents which nodes are infected after 1 step. Multiplying the start vector I0 with $A^{N}$ calculates which nodes are infected in N steps: each non zero entry represents an node infected after N steps. In the context of the FlipIt game, it is safe to assume that once a node is infected, it stays infected until the defender Flips the node. This means that after d steps, it is assumed that all nodes that could be reached in less than d steps from the node where the worm was dropped, are infected. In order to know how many nodes are infected after (for example) at most 3 steps, we have to consider nodes that are infected initially (step 1), or after 2 steps, or after 3 steps.  Calculating the sum of the three matrices $(A + A^{2} + A^{3}) $ results in a matrix that indicates for each node, the number of paths of length 1, 2 or 3 from i to j. Multiplying the start vector with this matrix, results in a vector that indicates which nodes will be infected in at most 3 steps . This technique can be applied to calculate the state of the network after any number of steps. 
Determining the length of the delay boils down to determining which configuration of infected nodes is considered as corresponding to the attacker having flipped the resource. It may be that all nodes need to be flipped, a sufficient amount of nodes, or a (set of) particular nodes.
--Hier nog een beetje verder uitschrijven hoe je exact de d dan moet bepalen, gegeven dat je niet weet in welke knoop het virus gedropt wordt--

### 2015-06-22 doc/template/chap-9.tex
nelezen verbeteringen

denoted as an infection-delay variable \textit{d} (called 'delay' for short in the remainder of this paper). If we want to measure how long it takes for the virus to

@@

\section{Playing periodically with propagation delay}
\label{ch2:periodicvirus}

\subsection{Formalization the benefit formula including the infection-delay}

>>>>

denoted as an propagation delay variable \textit{d} (called 'delay' for short in the remainder of this paper). If we want to measure how long it takes for the virus to

@@

\section{Formalization of the periodic game with propagation delay}
\label{ch2:periodicvirus}
